Erosion of somatic tissue identity with loss of the X-linked intellectual disability factor KDM5C

### 2024-08-26 submission/manuscript.Rmd
kdm5c binding at WT vs KO DEGs
draft wgbs figure

## KDM5C promotes _de novo_ DNA methylation at germline genes harboring CpG islands

@@

WTvsKO_hypo <- WTvsKO[WTvsKO$meth.diff <= -25, ]

@@

We then assessed differentially methylated promoters in wild-type versus _Kdm5c_-KO exEpiLCs and found `r nrow(WTvsKO_hypo)` germline gene promtoers were significantly hypomethylated with _Kdm5c_ loss (Figure 6I). Hypomethylated promoters included genes consistently dysregulated in _Kdm5c_-KO cells, such as _D1Pas1_ (methlyation difference = `r signif(getWGBSinfo(WTvsKO_hypo, "D1Pas1", "meth.diff"),4)`%, q-value = `r signif(getWGBSinfo(WTvsKO_hypo, "D1Pas1", "qvalue"),3)`) (Figure 6J).

>>>>

## KDM5C is recruited to CpG islands and facilitates _de novo_ DNA methylation at germline gene promoters

@@

WTvsKO_hypo <- WTvsKO[WTvsKO$meth.diff <= -25, ]
WTvsKO_hypo_5C <- WTvsKO_hypo[WTvsKO_hypo$KDM5C_binding == "Bound", ]

@@

We then assessed differentially methylated regions (DMRs) at germline gene promoters in wild-type versus _Kdm5c_-KO exEpiLCs. We found `r nrow(WTvsKO_hypo)` germline promoters were significantly hypomethylated with _Kdm5c_ loss (Figure 6I). About half of promoter DMRs hypomethylated in _Kdm5c_-KO exEpiLCs are direct targets of KDM5C in EpiLCs (`r nrow(WTvsKO_hypo_5C)` out of `r nrow(WTvsKO_hypo)` hypomethylated DMRs). Hypomethylated promoters included genes consistently dysregulated in _Kdm5c_-KO cells, such as _D1Pas1_ (methlyation difference = `r signif(getWGBSinfo(WTvsKO_hypo, "D1Pas1", "meth.diff"),4)`%, q-value = `r signif(getWGBSinfo(WTvsKO_hypo, "D1Pas1", "qvalue"),3)`) (Figure 6J).
